---
title: "Getting started with seqbench"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with seqbench}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 4)
```

`seqbench` answers one question: *when can I stop comparing algorithm A with algorithm B
without invalidating the inference?* It treats the benchmark as a sequential experiment on
**paired losses** (same instance, same seed) and maintains a confidence sequence for the mean
paired difference that stays valid under optional stopping, repeated inspection and adaptive
budgets. A pre-registered margin of practical equivalence `delta` turns the interval into one
of three declarations, or a decision to continue, or an inconclusive outcome when the budget
runs out.

## 1. The design comes first

Everything the procedure needs is fixed before any losses are observed:

```{r}
library(seqbench)

design <- comparison_design(
  alpha   = 0.05,        # P(any false declaration, ever) <= alpha
  margin  = 0.02,        # practical-equivalence margin delta, in loss units
  bounds  = c(0, 1),     # known bounds of the losses of both algorithms
  boundary = "betting",  # Waudby-Smith & Ramdas (2024); the default
  budget  = 800,         # total cost after which the outcome is "inconclusive"
  cost_per_round = 1     # cost of one (instance, seed) evaluation of both algorithms
)
design
```

The three hypotheses partition the line: A is relevantly better if `mu < -delta`, B if
`mu > delta`, and the two are practically equivalent if `|mu| <= delta`, where `mu` is the
mean of `loss_a - loss_b` over the population of instances. The margin is a domain decision,
never tuned after looking at data.

Before collecting anything you can ask how many instances a given distance to the
threshold will need:

```{r}
planning_horizon(design, distance = 0.05)             # guaranteed (Hoeffding), very conservative
planning_horizon(design, distance = 0.05, sd = 0.10)  # variance-based approximation
```

Typical stopping times are about half of these values.

## 2. Feeding paired losses

Instances are the sequential unit. Each row is one `(instance, seed)` pair with the losses of
both algorithms; several seeds of the same instance are averaged into a single observation.
Here we simulate a population where `loss_a - loss_b` has mean 0.05, i.e. A loses 0.05
more than B on average, so B should be declared relevantly better:

```{r}
set.seed(42)
simulate_batch <- function(instances, mu = 0.05, sd = 0.10) {
  h <- sd * sqrt(3)
  la <- runif(length(instances), 0.4, 0.8)
  data.frame(instance = instances, seed = 1L,
             loss_a = la, loss_b = la - mu + runif(length(instances), -h, h))
}

state <- initialize_comparison(design)
state <- update_comparison(state, simulate_batch(1:50))
state
```

Updating is incremental. Feed more instances whenever they are available; the procedure stops
at the first instance that produces a declaration and refuses data after that, because the
pre-registered protocol ends there:

```{r}
state <- update_comparison(state, simulate_batch(51:400))
stopping_decision(state)
```

## 3. Reading the result

`comparison_report()` (or `summary()`) returns the full contract: estimand, estimate, current
confidence sequence, the assumptions the guarantee relies on, diagnostics, stopping reason,
resources consumed, the instances and seeds in the order observed, and versions.

```{r}
comparison_report(state)
```

The trajectory is a tibble, one row per instance:

```{r}
tidy(state)[c(1:3, nrow(tidy(state))), ]
glance(state)
```

and the confidence sequence can be plotted against the number of instances, with the
equivalence band:

```{r}
ggplot2::autoplot(state)
```

## 4. What the package refuses

The guarantee rests on four assumptions, and the package will not silently convert a
violation into a valid-looking result:

```{r, error = TRUE}
fresh <- initialize_comparison(design)
update_comparison(fresh, data.frame(instance = 1, loss_a = 1.2, loss_b = 0.5))   # outside bounds
update_comparison(fresh, data.frame(instance = 1, loss_a = NA,  loss_b = 0.5))   # missing
```

Adding new seeds to an instance whose losses have already been seen is adaptive reuse, not
i.i.d. sampling, so it is an error too; plan all seeds of an instance before observing any of
them. Unpaired designs (`paired = FALSE`) are not implemented and say so.

## 5. Inconclusive is an outcome

When the budget ends before a declaration, the result is `"inconclusive"`, reported with the
final interval and the reason. It is never turned into a decision:

```{r}
tight <- comparison_design(alpha = 0.05, margin = 0.02, bounds = c(0, 1), n_max = 30)
st <- update_comparison(initialize_comparison(tight), simulate_batch(1:30, mu = 0.03))
stopping_decision(st)
```

## 6. Boundaries

| boundary | source | use |
|---|---|---|
| `"betting"` | hedged capital process, Waudby-Smith & Ramdas (2024) | default; tightest in the package's experiments |
| `"empirical_bernstein"` | predictable-plug-in empirical Bernstein, same source | conservative reference |
| `"hoeffding"` | predictable-plug-in Hoeffding, same source | distribution-free reference; very conservative |
| `"naive_fixed"` | fixed-sample t interval recomputed at every step | **invalid**; negative control for experiments only |

All three valid boundaries assume bounded losses, i.i.d. instances, seeds fixed before their
losses are seen, and no instance reuse. `naive_fixed` carries a `valid = FALSE` flag in every
output.
